---
title: "Homework 1"
subtitle: "4375 Machine Learning with Dr. Mazidi"
author: "Timothy Choi"
date: "1/24/22"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

This homework has two parts:

* Part 1 uses R for data exploration
* Part 2 uses C++ for data exploration

---

This homework is worth 100 points, 50 points each for Part 1 and Part 2. 

---

# Part 1: RStudio Data Exploration

**Instructions:** Follow the instructions for the 10 parts below. If the step asks you to make an observation or comment, write your answer in the white space above the gray code box for that step. 

## Step 1: Load and explore the data

* load library MASS (install at console, not in code)
* load the Boston dataframe using data(Boston)
* use str() on the data
* type ?Boston at the console
* Write 2-3 sentences about the data set below

Your commentary here:
Most of the data looks to be quantitative. chas looks to be the only qualitative data because of it only being 1 or 0.

```{r}
# step 1 code
library(MASS)
data(Boston)
str(Boston)
```

## Step 2: More data exploration

Use R commands to:

* display the first few rows
* display the last two rows
* display row 5
* display the first few rows of column 1 by combining head() and using indexing
* display the column names

```{r}
# step 2 code
head(Boston)
tail(Boston, n = 2)
Boston[5]
head(Boston[ , 1])
names(Boston)
```

## Step 3: More data exploration

For the crime column, show:

* the mean
* the median
* the range

```{r}
# step 3 code
mean(Boston$crim)
median(Boston$crim)
range(Boston$crim)
```

## Step 4: Data visualization

Create a histogram of the crime column, with an appropriate main heading. In the space below, state your conclusions about the crime variable:

Your commentary here:
From the data presented it seems that there is little to no crime in Boston.


```{r}
# step 4 code
hist(Boston$crim, main = "Boston Crime Rate", xlab = "Crime rate")

```

## Step 5: Finding correlations

Use the cor() function to see if there is a correlation between crime and median home value. In the space below, write a sentence or two on what this value might mean. Also write about whether or not the crime column might be useful to predict median home value.

Your commentary here:
There does not seem to be much of a correlation between crime and median home value because the value, -0.3883046, is closer to 0 than 1 or -1.
```{r}
# step 5 code
cor(Boston)

```

## Step 6: Finding potential correlations

Create a plot showing the median value on the y axis and number of rooms on the x axis. Create appropriate main, x and y labels, change the point color and style. [Reference for plots(http://www.statmethods.net/advgraphs/parameters.html) 

Use the cor() function to quantify the correlation between these two variables. Write a sentence or two summarizing what the graph and correlation tell you about these 2 variables.

Your commentary here:
There is a positive correlation between the 2 variables: number of rooms and median value. The more rooms there are the higher the median value is. 
```{r}
# step 6 code
plot(Boston$rm, Boston$medv, pch = 15, col ="blue", main = "Number of Rooms vs. Median Value", xlab = "Number of Rooms", ylab = "Median Value")
cor(Boston$rm, Boston$medv)
```

## Step 7: Evaluating potential predictors

Use R functions to determine if variable chas is a factor. Plot median value on the y axis and chas on the x axis. Make chas a factor and plot again.

Comment on the difference in meaning of the two graphs. Look back the description of the Boston data set you got with the ?Boston command to interpret the meaning of 0 and 1.  

Your commentary here:
The first graph does not really give much information other than that it shows the data lines on 0 or 1 on the x-axis. If you look at the 2nd graph it shows where the range of the values are with the circles being out liars. The box shows us where most of the data is located and the solid black line shows where the median value is.

```{r}
# step 7 code
is.factor(Boston$chas)
plot(Boston$chas, Boston$medv, main = "Chas vs Medium Value", xlab = "Chas", ylab = " Medium Value")
Boston$chas <- factor(Boston$chas, levels=c(0,1))
plot(Boston$chas, Boston$medv, main = "Chas vs Medium Value", xlab = " Chas", ylab = "Medium value")
```

## Step 8: Evaluating potential predictors

Explore the rad variable. What kind of variable is rad? What information do you get about this variable with the summary() function? Does the unique() function give you additional information? Use the sum() function to determine how many neighborhoods have rad equal to 24. Use R code to determine what percentage this is of the neighborhoods.

Your commentary here:
rad is an integer variable. With the summary function, you get the min, 1st quarter, median, mean, 3rd quarter, and max values. The unique function gives us every value of the data shown once which differs from what information was provided from the summary function. 

```{r}
# step 8 code
typeof(Boston$rad)
summary(Boston$rad)
unique(Boston$rad)
sum(Boston$rad == 24)
sum(Boston$rad == 24)/sum(!is.na(Boston$rad)) * 100
```

## Step 9: Adding a new potential predictor

Create a new variable called “far” using the ifelse() function that is TRUE if rad is 24 and FALSE otherwise. Make the variable a factor. Plot far and medv. What does the graph tell you? 

Your commentary here:
Houses that have a rad of 24 are on average lower cost than houses that do not have a rad value of 24.

```{r}
# step 9 code
far <- ifelse(Boston$rad == 24, TRUE, FALSE)
far <- factor(far)
plot(far, Boston$medv, main = "Far vs Medium value", xlab = "far", ylab = "Medium value")
```

## Step 10: Data exploration

* Create a summary of Boston just for columns 1, 6, 13 and 14 (crim, rm, lstat, medv)
* Use the which.max() function to find the neighborhood with the highest median value. See p. 176 in the pdf
* Display that row from the data set, but only columns 1, 6, 13 and 14
* Write a few sentences comparing this neighborhood and the city as a whole in terms of: crime, number of rooms, lower economic percent, median value.

Your commentary here:
The crime rate of the neighbor is half as much as the median. The number of rooms is about 1 less than than the maximum number of rooms.The neighborhood has the lowest value lstat in the city. Lastly, the neighborhood has the maximum value medimum value of the city.

```{r}
# step 10 code
summary(Boston[c(1,6,13,14)])
which.max(Boston$medv)
Boston[which.max(Boston$medv),c(1,6,13,14)]
```

# Part 2: C++

In this course we will get some experience writing machine learning algorithms from scratch in C++, and comparing performance to R. Part 2 of Homework 1 is designed to lay the foundation for writing custom machine learning algorithms in C++. 
	
To complete Part 2, first you will read in the Boston.csv file which just contains columns rm and medv. 
	
--- 

In the C++ IDE of your choice:
	
1	Read the csv file (now reduced to 2 columns) into 2 vectors of the appropriate type. See the reading in cpp picture in Piazza. 

2	Write the following functions: 

*	a function to find the sum of a numeric vector

*	a function to find the mean of a numeric vector

*	a function to find the median of a numeric vector

*	a function to find the range of a numeric vector

*	a function to compute covariance between rm and medv (see formula on p. 74 of pdf)

*	a function to compute correlation between rm and medv (see formula on p. 74 of pdf); Hint: sigma of a vector can be calculated as the square root of variance(v, v)

3	Call the functions described in a-d for rm and for medv. Call the covariance and correlation functions. Print results for each function.







